---
title: "GBM_Second_Project"
author: "Ali Kaynar"
date: "2024-01-27"
output: 
# rm(list=ls())
---
###(((((((.- .-.. .. / -.- .- -.-- -. .- .-.)))))))###
# Introduction
Briefly describe the purpose of the analysis and the dataset used.
Perform DEG analysis, Gene-Enrichment Analysis, and other statistical analysis.
# Setup
```{r setup, message=FALSE, warning=FALSE}
setwd("~/{PATH}") # Set your working directory
library(TCGAbiolinks)
library(DESeq2)
library(biomaRt)
library(clusterProfiler)
library(openxlsx)
library(xlsx)
library(writexl)
library(ggplot2)
library(dplyr)
#library(R.matlab)
```
# Data Loading
```{r data-loading}
# Load your datasets here
load("~/{PATH}/TCGA_GBM.Rdata") # TCGAbiolinks data structure
load("~/{PATH}/HPA_Info.RData") # HPA gene information
# Removed_samples (Duplicate) are "TCGA-06-0156-01A-03R-1849-01" and "TCGA-06-0211-01B-01R-1849-01"
```
Remove annotation version indicator for uniformity and compatibility with External Databases
Ensemble gene ID: "ENSG00000067177.15" --->> "ENSG00000067177". 
```{r echo=FALSE}
# 
TCGA_GBM@rowRanges@ranges@NAMES <- sapply(strsplit(TCGA_GBM@rowRanges@ranges@NAMES,"\\."), function(x) x[1])
TCGA_GBM@rowRanges@elementMetadata@listData[["gene_id"]] <- sapply(strsplit(TCGA_GBM@rowRanges@elementMetadata@listData[["gene_id"]],"\\."), function(x) x[1])
```
**Analysis includes Primary solid Tumor sample which includes only sample that have Tumor Purity Score(TPS) and Solid Tissue Normal. 
**Control vs Disease comparison and High TPS vs Low TPS comparison
Categorize the data: Solid Tissue Normal (NT), Recurrent Solid Tumor (TR) whic is not included in the study, Primary solid Tumor (TP), High Tumor Purity Score (high_tps) and Low Tumor Purity Score (low_tps)

Separation of data depending on definition. Later these separated structure could be combine. 
```{r}
TCGA_NT<-TCGA_GBM[,TCGA_GBM@colData@listData[["definition"]]=="Solid Tissue Normal"]

TCGA_TR<-TCGA_GBM[,TCGA_GBM@colData@listData[["definition"]]=="Recurrent Solid Tumor"]

TCGA_TP<-TCGA_GBM[,TCGA_GBM@colData@listData[["definition"]]=="Primary solid Tumor"]
TCGA_TP<-TCGA_TP[,-which(is.na(TCGA_TP@colData@listData[["paper_ABSOLUTE.purity"]]))]

# Extraction of High TPS and Low TPS samples.  First and fourth quantile used for separation. Samples sorted ascending purity score 
tps_col <- TCGA_TP@colData@listData[["paper_ABSOLUTE.purity"]]; tps_col
tps_quantile_25 <- quantile(tps_col, 0.25); tps_quantile_25
tps_quantile_75 <- quantile(tps_col, 0.75); tps_quantile_75

low_tps <- TCGA_TP[,tps_col <= tps_quantile_25]
mid_tps <- TCGA_TP[,tps_col > tps_quantile_25 & tps_col < tps_quantile_75]
high_tps <- TCGA_TP[,tps_col  >= tps_quantile_75]
```

# Analysis
Include analysis code chunks below

Comparison of NT and TR, which is not included in the study
```{r}
# Extraction of un-stranded gene count value from data structure  
NT_TR<-cbind(TCGA_NT,TCGA_TR) # Combine two data structure

NT_TR_clinic <- data.frame(barcode = NT_TR$barcode, definition = NT_TR$definition)
NT_TR_clinic$condition<-as.factor(NT_TR$definition) # Setting condition as factor

# Checking outlier!
# NOTE: TCGAanalyze_Preprocessing changes sample order! Be carefull.
NT_TR_count <- TCGAanalyze_Preprocessing(object = NT_TR,cor.cut = 0.80,datatype = "unstranded")

NT_TR_C <- NT_TR[which(rowMeans(NT_TR@assays@data@listData$unstranded)>=10),] ## keep expressed gene.
NT_TR_count<-data.matrix(NT_TR_C@assays@data@listData$unstranded)
colnames(NT_TR_count)<-colnames(NT_TR_C)
rownames(NT_TR_count)<-rownames(NT_TR_C)

# # Reaaragnge  sample order accordingly NT_TR datastuture
# NT_TR_count<-NT_TR_count[,NT_TR_clinic$barcode]
# 
# #Make sure matrix and clinic should be same order
# Cont_Samp_Ord<-rbind(colnames(NT_TR_count),NT_TR_clinic$barcode, NT_TR_C$barcode)
```

# DSEq2 Process
```{r warning=TRUE}
dds_NT_TR <- DESeqDataSetFromMatrix(countData=NT_TR_count,
                                       colData=NT_TR_clinic,
                                       design=~condition)
# Solid Tissue Normal set as reference
dds_NT_TR$condition <- relevel(dds_NT_TR$condition, ref = "Solid Tissue Normal") 

dds_NT_TR = DESeq(dds_NT_TR)
result_dds_NT_TR = results(dds_NT_TR)
summary(result_dds_NT_TR)
View(as.data.frame(result_dds_NT_TR))

nrow(result_dds_NT_TR)
sum( is.na(result_dds_NT_TR$pvalue) )
res_dds_NT_TR = result_dds_NT_TR[ ! is.na(result_dds_NT_TR$pvalue), ]
res_dds_NT_TR = res_dds_NT_TR[ order(res_dds_NT_TR$pvalue), ]

DESeq_Table_dds_NT_TR<-as.data.frame(res_dds_NT_TR)
```
Combining Gene information, including means expression and IDs,  with HPA gene informaton 
```{r}
TCGA_sum<-data.frame(Ensemble_ID=rownames(TCGA_GBM),
gene_name=TCGA_GBM@rowRanges@elementMetadata@listData[["gene_name"]],
hgnc_id=TCGA_GBM@rowRanges@elementMetadata@listData[["hgnc_id"]],
mean_NT=rowMeans(TCGA_NT@assays@data@listData$tpm_unstrand),
mean_TR=rowMeans(TCGA_TR@assays@data@listData$tpm_unstrand),
mean_TP=rowMeans(TCGA_TP@assays@data@listData$tpm_unstrand),
mean_low_tps=rowMeans(low_tps@assays@data@listData$tpm_unstrand),
mean_high_tps=rowMeans(high_tps@assays@data@listData$tpm_unstrand),
mean_mid_tps=rowMeans(mid_tps@assays@data@listData$tpm_unstrand))
rownames(TCGA_sum)<-TCGA_sum$Ensemble_ID

DESeq_Table_dds_NT_TR_2 <- merge(DESeq_Table_dds_NT_TR,TCGA_sum, by = 0, all.x = TRUE)
rownames(DESeq_Table_dds_NT_TR_2)<-DESeq_Table_dds_NT_TR_2$Row.names
DESeq_Table_dds_NT_TR_2 <- merge(DESeq_Table_dds_NT_TR_2,HPA_Info, by = 0, all.x = TRUE)

```
Gene Expression input for GEM generation. Mean of each genes calculated for all groups.
```{r}
Model_input_T<-TCGA_sum[,c("Ensemble_ID","mean_NT","mean_TR", "mean_TP","mean_low_tps","mean_mid_tps","mean_high_tps")]

# Create a "Results" folder in the current working directory
results_folder <- file.path("{PATH}", "Results")
if (!dir.exists(results_folder)) {
  dir.create(results_folder)
}
# Define the save path within the "Results" folder
save_path <- file.path(results_folder, "Model_input_T.txt")
# Save the table as a text file in the "Results" folder
write.table(Model_input_T, file = save_path, row.names = FALSE)
```
# Analysis
Include analysis code chunks below
Comparison of NT and TP

```{r}
# Extract un-stranded count value from data structure  
NT_TP<-cbind(TCGA_NT,TCGA_TP) # Combine two data structure

NT_TP_clinic <- data.frame(barcode = NT_TP$barcode, definition = NT_TP$definition)
NT_TP_clinic$condition<-as.factor(NT_TP$definition)

# Checking outlier!
# NOTE: TCGAanalyze_Preprocessing changes sample order!
NT_TP_count <- TCGAanalyze_Preprocessing(object = NT_TP,cor.cut = 0.80,datatype = "unstranded")

NT_TP_C <- NT_TP[which(rowMeans(NT_TP@assays@data@listData$unstranded)>=10),] ## keep expressed gene
NT_TP_count<-data.matrix(NT_TP_C@assays@data@listData$unstranded)
colnames(NT_TP_count)<-colnames(NT_TP_C)
rownames(NT_TP_count)<-rownames(NT_TP_C)

```

# DSEq2 Process
```{r warning=TRUE}
dds_NT_TP <- DESeqDataSetFromMatrix(countData=NT_TP_count,
                                    colData=NT_TP_clinic,
                                    design=~condition)
# Solid Tissue Normal set as reference
 dds_NT_TP$condition <- relevel(dds_NT_TP$condition, ref = "Solid Tissue Normal") 
# dds_NT_TP$condition <- relevel(dds_NT_TP$condition, ref = "Primary solid Tumor") # This code used to generate DEG results to generate NT GEM model compare to TP GEM model.

dds_NT_TP = DESeq(dds_NT_TP)
result_dds_NT_TP = results(dds_NT_TP)
summary(result_dds_NT_TP)
View(as.data.frame(result_dds_NT_TP))

nrow(result_dds_NT_TP)
sum( is.na(result_dds_NT_TP$pvalue) )
res_dds_NT_TP = result_dds_NT_TP[ ! is.na(result_dds_NT_TP$pvalue), ]
res_dds_NT_TP = res_dds_NT_TP[ order(res_dds_NT_TP$pvalue), ]

#DESeq_Table_dds_NT_TP<-as.data.frame(res_dds_NT_TP)

DESeq_Table_dds_TP_NT<-as.data.frame(res_dds_NT_TP)
```
Combining Gene information, including means expression and IDs,  with HPA gene informaton 
```{r}
DESeq_Table_dds_TP_NT_2 <- merge(DESeq_Table_dds_TP_NT,TCGA_sum, by = 0, all.x = TRUE)
rownames(DESeq_Table_dds_TP_NT_2)<-DESeq_Table_dds_TP_NT_2$Row.names
DESeq_Table_dds_TP_NT_2 <- merge(DESeq_Table_dds_TP_NT_2,HPA_Info, by = 0, all.x = TRUE)


```
# Analysis
Include analysis code chunks below

Comparison of high_tps and low_tps
```{r}
# Extract un-stranded count value from data structure  
high_low<-cbind(high_tps,low_tps) # Combine two data structure
array_high_low <- c(rep("high", dim(high_tps)[2]), rep("low", dim(low_tps)[2]))
high_low_clinic <- data.frame(barcode = high_low$barcode, definition = high_low$definition)
high_low_clinic$condition<-as.factor(array_high_low)

# Checking outlier!
# NOTE: TCGAanalyze_Preprocessing changes sample order!
high_low_count <- TCGAanalyze_Preprocessing(object = high_low,cor.cut = 0.80,datatype = "unstranded")

high_low_C <- high_low[which(rowMeans(high_low@assays@data@listData$unstranded)>=10),] ## keep expressed gene
high_low_count<-data.matrix(high_low_C@assays@data@listData$unstranded)
colnames(high_low_count)<-colnames(high_low_C)
rownames(high_low_count)<-rownames(high_low_C)

```

# DSEq2 Process
```{r warning=TRUE}
dds_high_low <- DESeqDataSetFromMatrix(countData=high_low_count,
                                       colData=high_low_clinic,
                                       design=~condition)
# Solid Tissue Normal set as reference
dds_high_low$condition <- relevel(dds_high_low$condition, ref = "low") 

dds_high_low = DESeq(dds_high_low)
result_dds_high_low = results(dds_high_low)
summary(result_dds_high_low)
View(as.data.frame(result_dds_high_low))

nrow(result_dds_high_low)
sum( is.na(result_dds_high_low$pvalue) )
res_dds_high_low = result_dds_high_low[ ! is.na(result_dds_high_low$pvalue), ]
res_dds_high_low = res_dds_high_low[ order(res_dds_high_low$pvalue), ]

DESeq_Table_dds_high_low<-as.data.frame(res_dds_high_low)
```
Combining Gene information, including means expression and IDs,  with HPA gene informaton 
```{r}
DESeq_Table_dds_high_low_2 <- merge(DESeq_Table_dds_high_low,TCGA_sum, by = 0, all.x = TRUE)
rownames(DESeq_Table_dds_high_low_2)<-DESeq_Table_dds_high_low_2$Row.names
DESeq_Table_dds_high_low_2 <- merge(DESeq_Table_dds_high_low_2,HPA_Info, by = 0, all.x = TRUE)

```
# Analysis
Include analysis code chunks below
Comparison of NT and low_tps
```{r}
# Extract un-stranded count value from data structure  
NT_low<-cbind(TCGA_NT,low_tps) # Combine two data structure
NT_low_clinic <- data.frame(barcode = NT_low$barcode, definition = NT_low$definition)
NT_low_clinic$condition<-as.factor(NT_low$definition)

# Checking outlier!
# NOTE: TCGAanalyze_Preprocessing changes sample order!
NT_low_count <- TCGAanalyze_Preprocessing(object = NT_low,cor.cut = 0.80,datatype = "unstranded")

NT_low_C <- NT_low[which(rowMeans(NT_low@assays@data@listData$unstranded)>=10),] ## keep expressed gene
NT_low_count<-data.matrix(NT_low_C@assays@data@listData$unstranded)
colnames(NT_low_count)<-colnames(NT_low_C)
rownames(NT_low_count)<-rownames(NT_low_C)

```

# DSEq2 Process
```{r warning=TRUE}
dds_NT_low <- DESeqDataSetFromMatrix(countData=NT_low_count,
                                       colData=NT_low_clinic,
                                       design=~condition)
# Solid Tissue Normal set as reference
dds_NT_low$condition <- relevel(dds_NT_low$condition, ref = "Solid Tissue Normal") 

dds_NT_low = DESeq(dds_NT_low)
result_dds_NT_low = results(dds_NT_low)
summary(result_dds_NT_low)
View(as.data.frame(result_dds_NT_low))

nrow(result_dds_NT_low)
sum( is.na(result_dds_NT_low$pvalue) )
res_dds_NT_low = result_dds_NT_low[ ! is.na(result_dds_NT_low$pvalue), ]
res_dds_NT_low = res_dds_NT_low[ order(res_dds_NT_low$pvalue), ]

DESeq_Table_dds_NT_low<-as.data.frame(res_dds_NT_low)
```
Combining Gene information, including means expression and IDs,  with HPA gene informaton 
```{r}
DESeq_Table_dds_NT_low_2 <- merge(DESeq_Table_dds_NT_low,TCGA_sum, by = 0, all.x = TRUE)
rownames(DESeq_Table_dds_NT_low_2)<-DESeq_Table_dds_NT_low_2$Row.names
DESeq_Table_dds_NT_low_2 <- merge(DESeq_Table_dds_NT_low_2,HPA_Info, by = 0, all.x = TRUE)

```
# Analysis
Include analysis code chunks below
Comparison of NT and high_tps
```{r}
# Extract un-stranded count value from data structure  
NT_high<-cbind(TCGA_NT,high_tps) # Combine two data structure
NT_high_clinic <- data.frame(barcode = NT_high$barcode, definition = NT_high$definition)
NT_high_clinic$condition<-as.factor(NT_high$definition)

# Checking outlier!
# NOTE: TCGAanalyze_Preprocessing changes sample order!
NT_high_count <- TCGAanalyze_Preprocessing(object = NT_high,cor.cut = 0.80,datatype = "unstranded")

NT_high_C <- NT_high[which(rowMeans(NT_high@assays@data@listData$unstranded)>=10),] ## keep expressed gene
NT_high_count<-data.matrix(NT_high_C@assays@data@listData$unstranded)
colnames(NT_high_count)<-colnames(NT_high_C)
rownames(NT_high_count)<-rownames(NT_high_C)

```

# DSEq2 Process
```{r warning=TRUE}
dds_NT_high <- DESeqDataSetFromMatrix(countData=NT_high_count,
                                     colData=NT_high_clinic,
                                     design=~condition)
# Solid Tissue Normal set as reference
dds_NT_high$condition <- relevel(dds_NT_high$condition, ref = "Solid Tissue Normal") 

dds_NT_high = DESeq(dds_NT_high)
result_dds_NT_high = results(dds_NT_high)
summary(result_dds_NT_high)
View(as.data.frame(result_dds_NT_high))

nrow(result_dds_NT_high)
sum( is.na(result_dds_NT_high$pvalue) )
res_dds_NT_high = result_dds_NT_high[ ! is.na(result_dds_NT_high$pvalue), ]
res_dds_NT_high = res_dds_NT_high[ order(res_dds_NT_high$pvalue), ]

DESeq_Table_dds_NT_high<-as.data.frame(res_dds_NT_high)
```
Combining Gene information, including means expression and IDs,  with HPA gene informaton 
```{r}
DESeq_Table_dds_NT_high_2 <- merge(DESeq_Table_dds_NT_high,TCGA_sum, by = 0, all.x = TRUE)
rownames(DESeq_Table_dds_NT_high_2)<-DESeq_Table_dds_NT_high_2$Row.names
DESeq_Table_dds_NT_high_2 <- merge(DESeq_Table_dds_NT_high_2,HPA_Info, by = 0, all.x = TRUE)

```

Analysis of Purity and Gene Expression data

```{r}
Pur_1<- as.numeric(TCGA_TP@colData@listData[["paper_ABSOLUTE.purity"]])
```

```{r}
# Shapiro-Wilk test for normality
shapiro.test(Pur_1)
#p-value indicates significance, we reject the null hypothesis that the data is normally distributed.

```

# ENRICHMENT ANALYSIS
Required packages
```{r}
library(clusterProfiler)
library(dplyr)
library(tidyr)
library(DOSE)
library(GO.db)
library(org.Hs.eg.db)
library(GSEABase)
library(viridis)  # cividis color palette
source('/{PATH}/deg_GoTerm_clusterProfiler.R') # it makes the result as a table 
```

Volcano plot

```{r pressure, echo=FALSE}
# Create volcano plot using the DEG result (RES_DEG). This chunk is automated to create figures and saving it-naming it.
Comparison<- 'High TPS vs Low TPS'; # 
# Save the res
RES_DEG<-DESeq_Table_dds_high_low_2

rownames(RES_DEG)<-RES_DEG$Ensemble_ID
deg_results_df1<- subset(RES_DEG, padj < 1e-10 & abs(log2FoldChange) > 1) # We focused on P-value rather than fold change.
deg_results_df<- RES_DEG[,c("Ensemble_ID","gene_name", "baseMean","log2FoldChange", "lfcSE","stat","pvalue","padj")]

# Define the file path for saving the Excel file
file_path <- paste0(results_folder, "/", Comparison,".xlsx")
# Write the top_genes dataframe to an Excel file
write_xlsx(deg_results_df1, file_path)

deg_results_df <- deg_results_df %>% arrange(padj)

volcano_plot <- ggplot(data = deg_results_df, aes(x = log2FoldChange, y = -log10(padj))) +
  geom_point(size = 1, color = ifelse(deg_results_df$padj < 1e-10 & abs(deg_results_df$log2FoldChange) > 1, "red", "black")) +
  geom_hline(yintercept = -log10(1e-10), linetype = "dashed", color = "blue") + #horizontal line
  geom_vline(xintercept = c(-1, 1), linetype = "dashed", color = "blue") + # vertical line
  labs(x = "log2FoldChange", y = "-log10(padj-value)", title = paste0("Volcano Plot Highlighting Significant Changes in ", Comparison)) + #  modifying the labels and title 
  theme_minimal() + # make it simple: removes grid lines, panel borders, colors
  theme(text = element_text(size = 12),  # base font size
        axis.title = element_text(size = rel(2)),  # axis title font size
        axis.text = element_text(size = rel(1.5)),  # axis text font size
        plot.title = element_text(size = rel(2.5)))  # plot title font size

# Add text annotations for the top 50 variable names
top_genes <- subset(deg_results_df, padj < 1e-10 & abs(log2FoldChange) > 1)[1:50, ]
volcano_plot <- volcano_plot +
  geom_text(data = top_genes, aes(label = gene_name), vjust = -0.5, size = 6)  # text size adjustment

volcano_plot

# Save the volcano plot as tiff image file with white background
ggsave(filename = paste0(results_folder, "/", "Volcano Plot Highlighting Significant Changes in", Comparison,".tiff"),
       plot = volcano_plot,
       width = 16,
       height = 12,
       dpi = 400,
       bg = "white")
```

Go Term enrichment
```{r}
# Interested gene list (DEG significant genes)
## mart = useMart("ensembl",dataset = "hsapiens_gene_ensembl")
# mart <- useMart("ensembl", host = "https://www.ensembl.org", dataset = "hsapiens_gene_ensembl")
# Comparison<- 'NT vs TP';
ont1 <- "CC"# "BP" "MF"   
# CC cellular component
# MF: Molecular Function 
# BP: Biological Process
header_F <- paste0("Highlighting Significant GO Terms ",ont1, "; ", Comparison, sep="")
# Perform GO enrichment
background <- as.matrix(deg_results_df$gene_name)
int_gene <- as.matrix(deg_results_df1$gene_name) # Check if conversion to matrix is necessary; typically, a vector is used
pathway <- enrichGO(gene = int_gene, OrgDb = 'org.Hs.eg.db', keyType = "SYMBOL", ont = ont1, universe = background, pAdjustMethod = "BH", qvalueCutoff = 0.05) # Output

# GeneExp<-NULL
# for (i in 1:length(pathway@result[["BgRatio"]])){
#   gene_ratio <- pathway@result[["GeneRatio"]][i]
#   s_gene_ratio <- as.numeric(strsplit(gene_ratio, "/")[[1]])
#   Bg_ratio <- pathway@result[["BgRatio"]][i]
#   s_Bg_ratio <- as.numeric(strsplit(Bg_ratio, "/")[[1]])
#   GeneExp[i]<-(s_Bg_ratio[1]*s_gene_ratio[2])/s_Bg_ratio[2]
# }
# pathway@result[["GeneExp"]]<-GeneExp

# save the table
path_table_all<-deg_GoTerm_clusterProfiler(pathway)
write.table(pathway, file = paste0(results_folder, "/",header_F,".txt"), sep = "\t", row.names = FALSE, col.names = TRUE, quote = FALSE)
```

```{r pressure, echo=FALSE}
# path_table_all contains 'term_name', 'Count', and 'p.adjust. it's sorted and filtered to top 20 by p.adjust
top20 <- head(path_table_all[order(path_table_all$p.adjust), ], 20)

# The bar plot generate with ggplot, ordering bars based on p.adjust value
ggplot(top20, aes(x = reorder(term_name, -p.adjust), y = Count, fill = -p.adjust)) +
  geom_bar(stat = "identity") +
  coord_flip() +  # Flip coordinates to make bar plot horizontal 
  #scale_fill_gradient(low = "yellow", high = "darkblue", name = "Adjusted P-Value") +  # Color-blind friendly gradient
  scale_fill_viridis_c(option = "cividis",direction = -1, name = "Adjusted P-Value") +  # cividis color map was constructed according to p.adjust
  labs(title = header_F, x = "Term Name", y = "Count") + # labels
  theme_minimal(base_family = "sans") + # all font family will be sans-serif font 
  theme(
    plot.background = element_rect(fill = "white", colour = NA),  # plot background is white
    panel.background = element_rect(fill = "white", colour = NA),  # panel background is white
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"), # face = title
    legend.position = "right" # legend position
  ) +
  theme(plot.margin = margin(1, 1, 1.5, 1.5, "cm")) # margins on the top, right, bottom, and left. That ensure you can fit the term name in the figure.

# Save the plot
ggsave(paste0(results_folder, "/",header_F, ".tiff"), width = 12, height = 12, units = 'in', dpi = 300)
```

KEGG Pathway Analysis
```{r}
int_gene<-as.matrix(deg_results_df1$gene_name) # interested genes: significantly altered genes

# in KEGG pathway analysis entrez gene ID recommended and widely used. So convert gene symbol or ensemble gene id to entrez gene id
int_gene_kegg = getBM(attributes = c("hgnc_symbol","entrezgene_id"),values =int_gene ,filters = "hgnc_symbol",mart = mart)
int_gene_kegg<-int_gene_kegg %>% drop_na()

# In general your gene list after low expressed gene filter is background genes = universe.
background2 = getBM(attributes = c("hgnc_symbol","entrezgene_id"),values =deg_results_df$gene_name,filters = "hgnc_symbol",mart = mart)
background2<-background2 %>% drop_na()

#KEGG pathway analysis
ekegg = enrichKEGG(gene = as.character(int_gene_kegg$entrezgene_id),
                   organism = "hsa",
                   keyType = "kegg",
                   pvalueCutoff = 0.05,
                   pAdjustMethod = "BH",
                   universe=as.character(background2$entrezgene_id),
                   qvalueCutoff = 0.05,
                   use_internal_data = TRUE)

write.table(ekegg, file = paste0(results_folder, "/","Highlighting Significant KEGG Pathways; ", Comparison,".txt"), sep = "\t", row.names = FALSE, col.names = TRUE, quote = FALSE)
# some time significant term number lower that first 10 or 20 term yo wish to visualize. So it is better to check number of significant KEGG term. 
if (class(ekegg) == "enrichResult") {
  ekegg_df <- as.data.frame(ekegg)[order(as.data.frame(ekegg)$p.adjust), ]
  ekegg_df <- head(ekegg_df, 20)
} else {
  ekegg_df <- ekegg
}

# The bar plot generate with ggplot, ordering bars based on p.adjust value
ggplot(ekegg_df, aes(x = reorder(Description, -p.adjust), y = Count, fill = -p.adjust)) +
  geom_bar(stat = "identity") +
  coord_flip() +  # make it horizontal
  scale_fill_viridis_c(option = "cividis", direction = -1, name = "Adjusted P-Value") + 
  labs(title = paste("Highlighting Significant KEGG Pathways; ", Comparison), x = "Pathway", y = "Count") +
  theme_minimal(base_family = "sans") +
  theme(
    plot.background = element_rect(fill = "white", colour = NA),  
    panel.background = element_rect(fill = "white", colour = NA),  
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    legend.position = "right"
  ) +
  theme(plot.margin = margin(1, 1, 1.5, 1.5, "cm"))

# Save the plot
ggsave(paste0(results_folder, "/","Highlighting Significant KEGG Pathways; ", Comparison, ".tiff"), width = 12, height = 12, units = 'in', dpi = 300)

```


Saving DEG Results for GEM analysis to perform reporter metabolites analysis.
```{r}
library(writexl)
all_objects <- ls()

# filter object names based on that start with "DESeq" and doesn't end with "_2"
selected_names <- grep("^DESeq(?!.*_2$).*", all_objects, perl = TRUE, value = TRUE)
dataList <- sapply(selected_names, get, simplify = FALSE)
names(dataList) <- selected_names  # Set the names of dataList elements

excelFilePath <- "{PATH}/DEG_RES_2.xlsx"
write_xlsx(dataList, path = excelFilePath)


# filter object names based on that start with "DESeq_Table_dds_" and end with "_2"
selected_names_2 <- grep("^DESeq_Table_dds_.*_2$", all_objects, value = TRUE)

dataList_2 <- sapply(selected_names_2, get, simplify = FALSE) # Create a list of these data frames

filteredDataList_2 <- lapply(dataList_2, function(df) {
  df_filtered <- df[df$padj < 0.01, ]  # filter for rows where p.adjust < 0.01
  return(df_filtered)
})

excelFilePath <- "{PATH}/DEG_RES_4.xlsx"
names(filteredDataList_2) <- selected_names_2
# Now write the named list to the Excel file
write_xlsx(filteredDataList_2, path = excelFilePath)

```
